# Week 0: Discussion Section

### 1. Get Started
```{r}
packages <- c("here", "janitor", "tidyverse", "sf", "terra", "tmap", "spData", "spDataLarge", "geodata", "kableExtra", "viridisLite")
installed_packages <- packages %in% rownames(installed.packages())

if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}
```

Load packages
```{r}
library(here)
library(janitor)
library(tidyverse)
library(sf)
library(kableExtra)
```

### 2. Load Data
```{r}
gdw_df <- read_csv("data/gdw.csv") |> # load in the csv file
    clean_names() # standardize the names to look uniform
```

### 3. Data Exploration

The first 10 rows
```{r}
head(gdw_df,10) |> 
    kable()
```

The last 10 rows
```{r}
tail(gdw_df,10) |> 
    kable() |> 
    kable_styling()
```

Find the number of rows and colums
```{r}
# glimpse(gdw_df)
print(paste("# of rows:", nrow(gdw_df)))
print(paste("# of cols:", ncol(gdw_df)))
```

Print the column names
```{r}
names(gdw_df)
```

### 4. Index, Summarize, Subset Data
Extract country name
```{r}
countries_df <- gdw_df[,"country"]
countries_df
country_vec <- gdw_df[["country"]]
```

Find number of dams by dam type
```{r}
gdw_df |> 
  group_by(dam_type) |>
  summarise(count = n()) |>
  ungroup()
```

subset for damns only
```{r}
sub_dam <- gdw_df |>
  filter(dam_type == "Dam")
sub_dam
```

Re-order gdw_df by ascending order of installation year
```{r}
gdw_df <- gdw_df |> 
    arrange(year_dam)
```

### 5. Data Visualization
Make a bar graph of average height of dam/barrier in meters (dam_hgt_m) by country
```{r}
ggplot(
    data = gdw_df,
    mapping = aes(
        y = mean(dam_hgt_m),
        x = country
    )
) +
    geom_bar()
```